---
title: "Machine Learning Project"
author: "Rodrigo Peña"
date: "7/30/2022"
output: html_document
---

```{r setoptions, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE,cache=FALSE,message = FALSE)
```

# Synopsis

One thing that people regularly do is quantify how  much of a particular activity they do, but they rarely quantify how well they do it.The goal of this project is to predict the manner in which they did the exercise.
The main strategy that will be followed is: after cleaning the data, three subsets will be created: one for training, one for testing the different models, and a last one for validation of the chosen model.

# Data preparation

```{r Data}
if (file.exists('preddata.csv') == FALSE){
        download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv',
                      destfile = './preddata.csv',method = 'curl')     
}
if (file.exists('quiz.csv') == FALSE) {
        download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv',
                      destfile = './quiz.csv',method = 'curl')
}

preddata <- read.csv('preddata.csv')
quiz <- read.csv('quiz.csv')
library(caret)
library(rattle)
library(ElemStatLearn)
library(randomForest)
library(e1071)
library(h2o)
set.seed(68490)
```

We then proceed to modify the class of certain variables into factor and POSIXct.

```{r factor}
preddata$classe <- factor(preddata$classe)
preddata$user_name <- factor(preddata$user_name)
preddata$cvtd_timestamp <- as.POSIXct(preddata$cvtd_timestamp,format='%d/%m/%Y %H:%M')
```

We drop the variable X, which merely indicates the number of the observation. Likewise, certain variables that happen to have a considerable number of NA are removed.

```{r nas}
preddata <- preddata[,2:dim(preddata)[2]]
navles <- apply(preddata,2,function(x) any(is.na(x)))
preddata <- preddata[,!navles]
```

Now we create the data subsets.

```{r sets}
inTrain <- createDataPartition(y=preddata$classe,p=.6,list=FALSE)
training <- preddata[inTrain,]
testing <- preddata[-inTrain,]
inTrain <- createDataPartition(y=testing$classe,list=FALSE)
validation <- preddata[inTrain,]
testing <- preddata[-inTrain,]
```

Finally, we remove those variables that have a strikingly low variance, and that therefore have barely no prediction power. For this purpose we use the training subset.

```{r var}
novar <- nearZeroVar(training)
training <- training[,-novar]
testing <- testing[,-novar]
validation <- validation[,-novar]
```

# Model prediction

## Decision Tree

```{r tree}
modtree <- train(classe~.,method='rpart',data=training)
fancyRpartPlot(modtree$finalModel)
confusionMatrix(testing$classe,predict(modtree,testing))
```

## Support Vector Machine

```{r svm}
modsvm <- svm(classe~.,data=training)
confusionMatrix(testing$classe,predict(modsvm,testing))
```

## Gradient Boosted Trees

```{r gbm1, results=FALSE}
h2o.init()
predictors <- colnames(training)[1:57]
response <- colnames(training)[58]
trainsp <- as.h2o(training)
testsp <- as.h2o(testing)
modgbm <- h2o.gbm(x=predictors,y=response,training_frame = trainsp,
                  validation_frame = testsp,ntrees=100,distribution='multinomial')
predgbm <- predict(modgbm,testsp)
perf <- h2o.performance(modgbm,testsp)
```

```{r gbm2}
h2o.confusionMatrix(modgbm,testsp)
1-0.0008 #Accuracy
```

## Random Forest

```{r rf}
modrf <- randomForest(classe~.,data=training)
varImpPlot(modrf)
confusionMatrix(testing$classe,predict(modrf,testing))
```

# Chosen Model And Validation

The model that performs the best, in terms of accuracy, is the random forest. We proceed then to apply it to the validation subset.

```{r val}
conf<-confusionMatrix(validation$classe,predict(modrf,validation))
ac <- conf$overall[1]
conf
```

Using validation data, the resulting accuracy of the random forest model is `r ac`, which indicates a really precise prediction.
The associated expected out of sample error is, hence, `r 1-ac`.
